---
title: "Demo: renv + conda + reticulate"
format: html
---

# renv: quick interactive demo

This demo shows minimal commands to initialize renv and install packages, then demonstrates Python usage and reticulate.

## R / renv section

```{r}
# Initialize renv for the project (run once)
if (!requireNamespace("renv", quietly = TRUE)) install.packages("renv")
# renv::init()  # uncomment to initialize in a live session

# Example R code (will run if packages are available)
library(ggplot2)
library(dplyr)

# small example dataset
x_vals <- rnorm(100)
df <- data.frame(x = x_vals, y = rnorm(100) + x_vals)

p <- ggplot(df, aes(x = x, y = y)) + geom_point() + geom_smooth(method = "lm")
print(p)

# After installing packages, run renv::snapshot() to record versions
# renv::snapshot()
```

---

## Python / conda section (Quarto Python engine)
```{r}
reticulate::py_config()

# import pandas
pd <- import("pandas")
df <- pd$DataFrame(list(x = c(1,2,3), y = c(4,5,6)))
print(df)
```


```{python}
# Example Python snippet (requires the conda env from environment.yml)
import pandas as pd
import matplotlib.pyplot as plt

pdf = pd.DataFrame({"x": [1,2,3,4], "y": [2,3,5,7]})
print(pdf.head())

plt.plot(pdf['x'], pdf['y'], marker='o')
plt.title('Simple Python plot')
plt.show()
```

---

## Reticulate: call Python from R

```{r}
library(reticulate)
# Point to the conda env (adjust name if you changed it)
# reticulate::use_condaenv("workshop-dependency", required = TRUE)

# Show Python configuration
print(reticulate::py_config())

# Import pandas and convert to R data.frame
pd <- import("pandas")
py_df <- pd$DataFrame(list(a = c(1,2,3), b = c(4,5,6)))
print(py_df)

# Convert to R
r_df <- py_to_r(py_df)
print(r_df)
```

---

## Notes for attendees

- If you only have R/RStudio, follow the mac/windows steps in the slides to install Miniconda and create the environment
- If you cannot run python chunks locally, focus on the renv portion — renv demonstrates reproducible R package management
