---
title: "Demo (R Markdown): renv + conda with reticulate"
output: html_document
---

# Overview

This R Markdown demo shows:
1. **renv** for R package management
2. **Reticulate** for Python execution (REQUIRED in R Markdown)
3. **R-to-Python interoperability**

## Setup: Load reticulate and configure Python

**Important:** In R Markdown, you MUST use reticulate to run Python chunks.

```{r setup}
library(reticulate)
# Point to the conda environment
use_condaenv("workshop-dependency", required = TRUE)

# Verify Python configuration
py_config()
```

---

## R / renv section

```{r}
# Initialize renv for the project (run once)
if (!requireNamespace("renv", quietly = TRUE)) install.packages("renv")
# renv::init()  # uncomment to initialize in a live session

# Example R code (will run if packages are available)
library(ggplot2)
library(dplyr)

# small example dataset
x_vals <- rnorm(100)
df <- data.frame(x = x_vals, y = rnorm(100) + x_vals)

p <- ggplot(df, aes(x = x, y = y)) + geom_point() + geom_smooth(method = "lm")
print(p)

# After installing packages, run renv::snapshot() to record versions
# renv::snapshot()
```

---

## Python section (via reticulate)

After the setup chunk above, Python chunks will work via reticulate:

```{python}
# Example Python snippet
import pandas as pd
import matplotlib.pyplot as plt

pdf = pd.DataFrame({"x": [1,2,3,4], "y": [2,3,5,7]})
print(pdf.head())

plt.plot(pdf['x'], pdf['y'], marker='o')
plt.title('Simple Python plot via reticulate')
plt.show()
```

---

## Reticulate: R ↔ Python interoperability

Share objects between R and Python:

```{r}
# Import pandas and create a DataFrame from R
pd <- import("pandas")
py_df <- pd$DataFrame(list(a = c(1,2,3), b = c(4,5,6)))
print(py_df)

# Convert to R data.frame
r_df <- py_to_r(py_df)
print(class(r_df))
print(r_df)
```

You can also access Python objects from R chunks:

```{python}
# Create a Python object
my_python_list = [10, 20, 30, 40, 50]
```

```{r}
# Access it from R using py$
print(py$my_python_list)
```

---

## Notes for attendees

- **R Markdown requires reticulate** for all Python execution
- Must call `use_condaenv()` in a setup chunk before Python chunks
- Python chunks are executed by reticulate, not Jupyter
- Objects can be shared: use `py$object_name` in R, `r.object_name` in Python
- To render: Open in RStudio and click "Knit" or run `rmarkdown::render("demo.Rmd")`
